---
title: "Hands-on Exercise 1"
author: "Shermainn"
---

# Getting Started

## Install and launch R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
# eval: false
# echo: false

ggplot(data = exam_data,
       aes(x = RACE)) + geom_bar() + coord_flip()
```

# Plotting a histogram

```{r}
ggplot(data=exam_data, aes(x = MATHS, fill = GENDER)) +
  geom_histogram(bins=20, 
         color="grey30") +
  ggtitle("Distribution of Maths scores")

```


# Plotting a boxplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

# Plotting a scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
             linewidth=0.5)
```

# Plotting Facets (Trellis plot)

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

# Plotting a line graph (English Vs Science Scores by class)

```{r}

exam_long <- exam_data %>%
  pivot_longer(cols = c(SCIENCE, ENGLISH),
               names_to = "Subject",
               values_to = "Score")

ggplot(data = exam_long, 
       aes(x = CLASS, y = Score, color = Subject, group = Subject)) +
  geom_line() +
  labs(title = "Science and English Scores by Class")

```