---
title: "Hands-on_Ex03"
---

# 1. Overview

1.  I will be learning how to create interactive data visualization by using functions provided by `ggiraph` and `plotlyr` packages.
2.  I will also be learning how to create animated data visualization by using `gganimate` and `plotly r` packages. In addition, I will be able to reshape data using `tidyr` package, and (ii) process, wrangle and transform data by using `dplyr` package.

# 1.1 Getting Started

## 1.1.1. Install & Launch R packages

Install and Launch the following R packages:

-   [**ggiraph**](https://davidgohel.github.io/ggiraph/) for making ‘ggplot’ graphics interactive.

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**DT**](https://rstudio.github.io/DT/) provides an R interface to the JavaScript library [DataTables](https://datatables.net/) that create interactive table on html page.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

-   [**patchwork**](https://patchwork.data-imaginist.com/) for combining multiple ggplot2 graphs into one figure.

```{r}

```

## 1.1.2. Import Data

The code chunk below `read_csv()` of **readr** package will import the `Exam_data.csv` and save as **`exam_data`** as a tibble data frame.

```{r}

```

# 1.2 Getting Started

## 1.2.1. Install and launch R packages

Install and Launch the following R packages:

-   [**plotly**](https://plotly.com/r/), R library for plotting interactive statistical graphs.

-   [**gganimate**](https://gganimate.com/), an ggplot extension for creating animated statistical graphs.

-   [**gifski**](https://cran.r-project.org/web/packages/gifski/index.html) converts video frames to GIF animations using pngquant’s fancy features for efficient cross-frame palettes and temporal dithering. It produces animated GIFs that use thousands of colors per frame.

-   [**gapminder**](https://cran.r-project.org/web/packages/gapminder/index.html): An excerpt of the data available at Gapminder.org. We just want to use its *country_colors* scheme.

-   [**tidyverse**](https://www.tidyverse.org/), a family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate, tidyverse)
```

## 1.2.2 Import Data

Import data worksheet from GlobalPopulation Excel workbook.

```{r}

```
