---
title: "Hands-on Exercise 2"
author: "Shermainn"
---

# 1. Overview

In this exercise, I will be learning on how to:

1.  Control the placement of annotation on a graph using functions provided in ggrepel package
2.  Create professional publication quality figures using functions provided in ggthemes and hrbrthemes packages
3.  Plot composite figures by combining ggplot2 graphs using patchwork package

# 2. Getting Started

## 2.1. Install and Load required libraries

Code chunk below is used to check if all 4 R packages (ggrepel, ggthemes, hrbrthemes & patchwork) are installed and loaded onto the environment.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## 2.2. Import data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# 3. ggrepel

To prevent overlap of text, replace [geom_text()]{.underline} with **geom_text_repel()** and [geom_label()]{.underline} with **geom_label_repel()**.

| ggrepel | Original Text | Replace with       |
|---------|---------------|--------------------|
|         | geom_text()   | geom_text_repel()  |
|         | geom_label()  | geom_label_repel() |

::: panel-tabset
### Default

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### With ggrepel

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```

### With ggrepel with overlaps

```{r}
ggplot(data = exam_data, 
       aes(x = MATHS, 
           y = ENGLISH)) +
  geom_point() +
  geom_smooth(method = lm, 
              linewidth = 0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold",
                   max.overlaps = nrow(exam_data)*0.2) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```
:::

# 4. ggthemes

ggthemes provide ggplot2 themes by replicating the look of plots by Edward Tufte and Stephen Few.

+-----------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Category                                                                                                              | Description                                                                                                                                                                                                                     |
+=======================================================================================================================+=================================================================================================================================================================================================================================+
| **ggplot2 Built-in Themes**                                                                                           | 8 built-in themes: `theme_gray()`, `theme_bw()`, `theme_classic()`, `theme_dark()`, `theme_light()`, `theme_linedraw()`, `theme_minimal()`, and `theme_void()`                                                                  |
+-----------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **ggthemes (Extension)**                                                                                              | Additional themes from the `ggthemes` package: `theme_wsj()`, `theme_tufte()`, `theme_stata()`, `theme_solid()`, `theme_fivethirtyeight()`, `theme_few()`, `theme_excel()`, `theme_economist()`, and `theme_seq_gradient_pal()` |
|                                                                                                                       |                                                                                                                                                                                                                                 |
| More examples found in [ggthemes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/ "ggthemes"). |                                                                                                                                                                                                                                 |
+-----------------------------------------------------------------------------------------------------------------------+---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

::: panel-tabset
### gg2plot (using theme_grey())

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### ggtheme package (using theme_economist())

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```
:::

# 5. hrbrthemes

This package provides a base theme that focuses on typographic elements including labels (axis label) and fonts (font size). In addition, the second goal centers around productivity for a production workflow.

::: panel-tabset
Default

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

With hrbrthemes

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")
```

From the code chunk above:

-   `axis_title_size` - increase font size of axis

-   `base_size` - increase default axis label

-   `grid` - remove x-axis gridlines

# 
:::

# 6. patchwork

It is used to combine separate ggplot2 graphs into a single figure. It also provides auto-tagging capabilities. This package has a general syntax that combines:

-   Two-column Layout using the plus sign (+)

-   Parenthesis () to create a subplot group

-   Two-row Layout using the division sign (/)

::: panel-tabset
### Default - creating multiple graphs separately using ggplot2

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

p1#

p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")

p2#

p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

p3#
```

### Patchwork (using + to combine 2 plots)

```{r}
p1 + p2
```

### Patchwork (using "/","\|","()" to combine 3 plots)

```{r}
(p1 / p2) | p3
```

### Patchwork (with tag)

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### Patchwork (with insert)

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### Combine patchwork and ggtheme

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```

# 
:::

# 7. Additional Plots

## 7.1. Boxplot

## 7.2. Violin Plot

## 7.3. Notched Box Plot

## 7.4. Dot Plot

## 7.5. Pie Chart and Bar Chart

## 7.6. Raincloud Plot

## 7.7. Diverging Stacked Bar Chart

## 7.8. Packed Bars
